MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-12-11 Rebuttal.tex

		In dataset 1, Figure~\ref{fig:angle_bundles}f, bundles 7,10 and 6 show  bundles which  though parallel differ in angle and bundle width.It is a common example of variations in the datasets which we extract accurately.

		During additional experiments, we did apply our technique to a different class of fibers where the individual fibers are clearly visible, namely the glass-fibers. We have added the results to the supplementary section. 
		\begin{figure}
			\centering
			\includegraphics[width=0.7\linewidth]{images_pvis/clustering.pdf}
			\caption{Fiber bundles extraction}
			\label{fig:angle_bundles}
		\end{figure}}
		\item {
			As mentioned in the paper, 
			``The length and the radius parameters for the cylinders of MetaTracts decide how coarse the fiber cylinders are. These are dependent on the underlying fiber characteristic and the weaving pattern. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.  
			Simpler geometry (example dataset 2) was experimentally found to handle larger cylinders better.''

			Reviewer 3 and  4, asked  about the effect of the length threshold ``$\eta$".  ``$\eta$" decides how quickly the hierarchical clustering converges. As mentioned in the paper,  experimentally 0.3 to 0.6 gave qualitatively similar results. Quantitatively in our  tests $1.2\% - 5\%$ of fibers (total number of fibers $\sim$10000)were removed, while quantitatively small, without removing these clusters are difficult to separate.
			}
			\item{
				We have added a limitations section to the paper. We provide the approximate times of MetaTracts computation of our datasets in the Experimental Results section of the paper. Unfortunately the full scale performance study was not possible due to the time constraints of the minor revision.}
\end{enumerate}

### 2014-12-11 chapters/MetaTracts.tex
rebuttal reviewed
spell check pass

\item{ Which fiber bundles are of the same type of yarn\added{?} (i.e. which bundles show similar sizes or diameters, which is the largest or smallest fiber bundle?\added{)}}

@@

We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of \replacedWith{bundles}{them}. Spatial context refers to how two or more bundles interact with each other. \replacedWith{For example, spatial context}{It provides} answers \added{to} questions such as: Are these bundles in contact at a particular position in the data? What is the relative orientations of the contacting bundles? 
\replacedWith{Answering}{Providing answers to} the above queries from the volume renderings of XCT datasets or from visual inspection of particular 2D slices is non-trivial, even for experts. We \replacedWith{detail}{present} and evaluate a method which uses visualization techniques to gain insight into our data.
We \replacedWith{advance and interpret}{interpret and advance} techniques from diffusion tensor imaging to extract and visualize geometric structures from \replacedWith{3D X-ray computed tomography}{XCT} data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \replacedWith{lower resolutions and thus larger areas}{\textit{complete unit cells}} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=\linewidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
	\vskip-0.2cm
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach.
\replacedWith
{
Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.
} 
{
Starting from XCT data, we first generate MetaTracts (see Sec.~\ref{sec:approach}). Then we cluster the MetaTracts to generate and visualize fiber bundles (see Sec.~\ref{subsec:fiber-bundles}). We discuss the visualizations we created for our domain experts in Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} finally summarizes the user evaluation of our method.
}

@@

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, similarity measures such as proximity between fibers are used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and \removed{choosing} a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers \replacedWith{that}{which} are close for certain distance and then diverge, are clustered separately. Brun et al.~\cite{Brun2004} use\removed{s} normalized cuts along with a pairwise distances measure computed using a 9\removed{-}D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{Brun2004,jonasson2005, ODonnell2007}.
Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique, Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}), can be applied to the problem of organizing fiber tracts data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber tracts. In the ideal case fiber tracts belonging to the same bundle must remain ``close" to each other in the lower dimensional space. Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of point wise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}. 
The agglomerative hierarchical clustering method~\cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results. 
There are a number of difficulties in hierarchical clustering. First, computing all pairs\added{'} distances for tracts to generate the distance matrix is time consuming~\cite{Garyfallidis2012}. Second, a ``correct'' distance measure to compare tracts must be chosen. Third, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

@@

\subsection {Visual Analysis and Modeling of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fiber extraction from high resolution data where the individual fibers are clearly discernible. Fritz et al.~\cite{Fritz2009} proposed interactive workflows%for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. %This approach allows analyzing fiber orientations based on direction transfer functions. 
Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites.%They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Lomov et al.~\cite{Lomov2010} discusses the problems and current available solutions in geometric modeling of three dimensional composites.
%Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position? The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.

@@

The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location \replacedWith{becomes}{gets} associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation. We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image.The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure. \replacedWith
{The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}
{The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}

@@

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute the same metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  

@@

First, vesselness in biology is computed for different scales because the vessels can be of different sizes. In our case, usually the width of individual fibers are known a priori.

@@

Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively. Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest. Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable. The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.

@@

\replacedWith
{We note some intrinsic differences between the DTI and our XCT data. Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}. The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding. This is not the case for our data. The local orientaion at each voxel is inherently more noisy. Thus, techniques popular in DTI cannot be directly applied to XCT data.}
{We also note here some intrinsic differences between the DTI and our XCT data. Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}. The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding. This is not the case for our data. The local orientation at each voxel is inherently more noisy.}

@@

\subsection {MetaTracts properties and description} 
\label{subsec:mtprop}
%\begin{figure*}
%  \centering
%  \subfloat[Meta tract generation in 2D]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1b.eps}}
%  \subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified examples of meta tract generation process and orientation based similarity measure}
%  \label{fig:algo}
%\end{figure*}
%\begin{figure*}
%  \centering
%  \subfloat[]{\includegraphics[width=0.2\textwidth]{imagesMetaTracts/algo-1d}}
%  %\subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified example of meta tract generation process.}
%  \label{fig:algo}
%\end{figure*}
The input to this step is a grid in $\mathbb{R}^3$ where each grid point is associated with a normalized local orientation and a corresponding $R_H$ value.
Traditional integral curve based techniques cannot be used to extract fiber bundle traces from the reliable Hessians because of the spurious nature of the Hessian based local orientation. Thus instead of building fiber traces, we wish to find an abstract representation of the fibers. We build on the key assumptions on the data, namely ``local orientation" and ``connectivity" while taking into account the noise and lack of resolution. We do this by interpreting the underlying geometric structure of the fibers as a set of \removed{polylines(cylinders)} cylinders.
 MetaTacts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data. MetaTracts share the following \textit{properties}:

\begin{enumerate}[noitemsep]
\item{\replacedWith{Each MetaTract is}{MetaTract are} associated with a continuous set of cylinders.}
\item{\replacedWith{Each MetaTract is}{MetaTract are} associated with a start point at a grid vertex.}
\item{Individual cylinders in a MetaTract have a constant length, a constant radius and a start point (which is also a grid vertex).}
\item{Individual cylinders in a MetaTract (except the first one) are connected to their previous cylinders at their start point.}
\item{Individual cylinders are parallel to the local orientation at their start points.}
\end{enumerate}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$ (the procedure trivially extends to $\mathbb{R}^3$). In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles. The reliable Hessian step generates a local orientation vector in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1]. In Figure~\ref{fig:algo}, regions with unreliable Hessian have been marked in blue. Figure~\ref{fig:algo}c shows \replacedWith{a single}{an individual} cylinder of a MetaTract. Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow. We generate a rectangle of length $L$ and radius $R$ (properties 3 and 5 in Section~\ref{subsec:mtprop}). 
%
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=\linewidth]{images_pvis/mt_gen.pdf}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (b) \replacedWith{MetaTract generation, region with $R_H$ less than threshold}{Generation process}. (a) \replacedWith{Close up of}{Zoom in of the white} marked region in (b). (c) \replacedWith{Parts of a single MetaTract}{Individual cylinder}.}
  \label{fig:algo}
    %\vspace{-1.5em}
    \vskip-0.2cm
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible \textit{``candidate vertices"} for the start point of the next cylinder in a particular MetaTract. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
Based on assumptions on our data (local orientation and connectivity), the order of the candidate vertices is based on the following \textit{characteristics}:

@@

\section {Fiber Bundle Generation}
\label{subsec:fiber-bundles}
\replacedWith{The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles.}
{The set of MetaTracts generated by the previous step needs to be clustered in order to extract the final fiber bundles. }
\replacedWith{Carbon fibers hav}{Carbon fibers show} both orientation and geometric proximity information, both of which can be used for clustering. Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with \replacedWith{fibers which partially overlap}{partially overlapping fibers} (Figure~\ref{fig:length_distribution}b). 
We also found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).  
\added{So} instead of creating a heuristic \replacedWith{which}{to} artificially combine the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity. The two step clustering also made the problem more tractable and the parameters intuitive for the end users. For the orientation clustering, we use dimension reduction followed by K-means clustering. We then use hierarchical clustering to further subdivide each cluster based on the geometric proximity measures.

@@

shortest tracts (tracts less than $\eta$ times the median length, set to 0.6) in each bundle. We then reapplied hierarchical clustering.

>>>>

\item{ Which fiber bundles are of the same type of yarn? I.e. which bundles show similar sizes or diameters, which is the largest or smallest fiber bundle\added{?}}

@@

We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of \replacedWith{bundles}{them}. Spatial context refers to how two or more bundles interact with each other. \replacedWith{For example, spatial context}{It provides} answers \added{to} questions such as: Are these bundles in contact at a particular position in the data? What are the relative orientations of the contacting bundles? 
\replacedWith{Answering}{Providing answers to} the above queries from the volume renderings of XCT datasets or from visual inspection of particular 2D slices is non-trivial, even for experts. We \replacedWith{detail}{present} and evaluate a method which uses visualization techniques to gain insight into our data.
We \replacedWith{advance and interpret}{interpret and advance} techniques from diffusion tensor imaging to extract and visualize geometric structures from \replacedWith{3D X-ray computed tomography}{XCT} data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \replacedWith{lower resolutions and thus larger areas}{\textit{complete unit cells}} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=\linewidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
	\vskip-0.2cm
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach.
\replacedWith
{
Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.
} 
{
Starting from XCT data, we first generate MetaTracts (see Sec.~\ref{sec:approach}). Then we cluster the MetaTracts to generate and visualize fiber bundles (see Sec.~\ref{subsec:fiber-bundles}). We discuss the visualizations we created for our domain experts in Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} finally summarizes the user evaluation of our method.
}

@@

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, similarity measures such as proximity between fibers are used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and \removed{choosing} a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the closest distances mean, so that fibers \replacedWith{that}{which} are close for certain distance and then diverge, are clustered separately. Brun et al.~\cite{Brun2004} use\removed{s} normalized cuts along with a pairwise distances measure computed using a 9\removed{-}D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{Brun2004,jonasson2005, ODonnell2007}.
Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique, Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}), can be applied to the problem of organizing fiber tracts data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber tracts. In the ideal case fiber tracts belonging to the same bundle must remain ``close" to each other in the lower dimensional space. Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of point wise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}. 
The agglomerative hierarchical clustering method~\cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results. 
There are a number of difficulties in hierarchical clustering. First, computing all pairs\added{'} distances for tracts to generate the distance matrix is time consuming~\cite{Garyfallidis2012}. Second, a ``correct'' distance measure to compare tracts must be chosen. Third, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

@@

\subsection {Visual Analysis and Modeling of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focuses on individual objects such as fiber extraction from high resolution data where the individual fibers are clearly discernible. Fritz et al.~\cite{Fritz2009} proposed interactive workflows. %for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. %This approach allows analyzing fiber orientations based on direction transfer functions. 
Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. %They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations.
Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Lomov et al.~\cite{Lomov2010} discusses the problems and current available solutions in geometric modeling of three dimensional composites.
%Modeling of the composites, first starts with establishing the topology of the structure, which translates to answering if a particular bundle is in contact with another at a particular position? The second step builds the geometry of the model, answering questions relating to placement of bundles in space their orientations and dimensions.

@@

The input to the first stage of our pipeline, is the original scalar dataset in a uniform lattice grid in $\mathbb{R}^3$. As output, each grid location \replacedWith{becomes}{gets} associated with an unit vector representing the local orientation and a real value [0,1] that represents a measure of the reliability of the local orientation. We approximate the local orientation by eigenvalue analysis of the Hessian matrix computed at each voxel.
The Hessian matrix represents the local curvature. The eigen decomposition of the Hessian matrix gives the eigenvectors which describe the local second-order structure of the image. The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure. \replacedWith
{The eigenvector corresponding to the smallest eigenvalue gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}
{The eigenvector corresponding to the smallest eigenvalue (also referred as the principal direction) gives the direction along which the curvature is smallest. This direction coincides with the direction of the tubular structure.}

@@

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute the same metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign ($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  

@@

First, vesselness in biology is computed for different scales because the vessels can be of different sizes. In our case, usually the widths of individual fibers are known a priori.

@@

Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively. Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest. Note, the dark regions within bundles are regions where the $R_H$ is less than the threshold and are unreliable. The bundles are not uniformly colored and the Hessians and the corresponding directions are noisy.

@@

\replacedWith
{We note some intrinsic differences between the DTI and our XCT data. Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}. The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding. This is not the case for our data. The local orientaion at each voxel is inherently more noisy. Thus, techniques popular in DTI cannot be directly applied to XCT data.}
{We also note here some intrinsic differences between the DTI and our XCT data. Fiber traces can be created in DTI using a standard fiber tracking algorithm following the principal direction of diffusion using a fourth order Runge-Kutta method~\cite{Brun2003}. The principal direction based on the Hessian matrix works best when the tubular structures are well separated from the surrounding. This is not the case for our data. The local orientations at each voxel are inherently noisier.}

@@

\subsection {MetaTracts properties and description} 
\label{subsec:mtprop}
%\begin{figure*}
%  \centering
%  \subfloat[Meta tract generation in 2D]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1b.eps}}
%  \subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified examples of meta tract generation process and orientation based similarity measure}
%  \label{fig:algo}
%\end{figure*}
%\begin{figure*}
%  \centering
%  \subfloat[]{\includegraphics[width=0.2\textwidth]{imagesMetaTracts/algo-1d}}
%  %\subfloat[Orientation based similarity measure]{\includegraphics[width=0.5\linewidth]{imagesMetaTracts/algo-1c.eps}}
%  \caption{Simplified example of meta tract generation process.}
%  \label{fig:algo}
%\end{figure*}
The input to this step is a grid in $\mathbb{R}^3$ where each grid point is associated with a normalized local orientation and a corresponding $R_H$ value.
Traditional integral curve based techniques cannot be used to extract fiber bundle traces from the reliable Hessians because of the spurious nature of the Hessian based local orientation. Thus instead of building fiber traces, we wish to find an abstract representation of the fibers. We build on the key assumptions on the data, namely ``local orientation" and ``connectivity" while taking into account the noise and lack of resolution. We do this by interpreting the underlying geometric structure of the fibers as a set of \removed{polylines(cylinders)} cylinders.
 MetaTracts  are a coarse and simple approximation of integral curves in the form of a continuous chain of cylindrical tubes in $\mathbb{R}^3$ traversing the fiber bundles embedded in the input data. MetaTracts share the following \textit{properties}:

\begin{enumerate}[noitemsep]
\item{\replacedWith{Each MetaTract is}{MetaTracts are} associated with a continuous set of cylinders.}
\item{\replacedWith{Each MetaTract is}{MetaTracts are} associated with a start point at a grid vertex.}
\item{Individual cylinders in MetaTracts have constant lengths, radii, and start points (which are also grid's vertices).}
\item{Individual cylinders in MetaTracts (except the first one) are connected to their previous cylinders at their start point.}
\item{Individual cylinders are parallel to the local orientation at their start points.}
\end{enumerate}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In this section we explain the process of MetaTracts generation in $\mathbb{R}^2$ (the procedure trivially extends to $\mathbb{R}^3$). In $\mathbb{R}^2$, all the above properties hold except that cylinders are replaced by rectangles. The reliable Hessian step generates a local orientation vector in $\mathbb{R}^2$ and a reliability measure $R_H$[0,1]. In Figure~\ref{fig:algo}, regions with unreliable Hessian have been marked in blue. Figure~\ref{fig:algo}c shows \replacedWith{a single}{an individual} cylinder of a MetaTract. Let the seed point associated with the MetaTract be grid point $C_{p}$ (property 2 in Sec.~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow. We generate a rectangle of length $L$ and radius $R$ (properties 3 and 5 in Section~\ref{subsec:mtprop}). 
%
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=\linewidth]{images_pvis/mt_gen.pdf}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (b) \replacedWith{MetaTract generation, region with $R_H$ less than threshold}{Generation process}. (a) \replacedWith{Close up of}{Zoom in of the white} marked region in (b). (c) \replacedWith{Parts of a single MetaTract}{Individual cylinder}.}
  \label{fig:algo}
    %\vspace{-1.5em}
    \vskip-0.2cm
\end{figure}
The set of vertices which are located in the green region but not in the blue one are potential \textit{``candidate vertices"} for the start point of the next cylinder in a MetaTract. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
Based on assumptions on our data (local orientation and connectivity), the order of the candidate vertices is based on the following \textit{characteristics}:

@@

\section {Fiber Bundle Generation}
\label{subsec:fiber-bundles}
\replacedWith{The output of the above step is a set of MetaTracts. The next step is to cluster the MetaTracts into bundles.}
{The set of MetaTracts generated by the previous step needs to be clustered in order to extract the final fiber bundles. }
\replacedWith{Carbon fibers hav}{Carbon fibers show} both orientation and geometric proximity information, both of which can be used for clustering. Experimentally we found that the orientation measure was very reliable, while measures of geometric proximity had problems with \replacedWith{fibers which partially overlap}{partially overlapping fibers} (Figure~\ref{fig:length_distribution}b). 
We also found that different clustering techniques performed preferably for different measures (Sec.~\ref{subsec:clus_choice}).  
\added{So} instead of creating a heuristic \replacedWith{which}{to} artificially combine the orientation and geometric proximity measures, we first cluster based on orientation and then further subdivide each orientation cluster using geometric proximity. The two-step clustering also made the problem more tractable and the parameters intuitive for the end users. For the orientation clustering, we use dimension reduction followed by K-means clustering. We then use hierarchical clustering to further subdivide each cluster based on the geometric proximity measures.

@@

shortest tracts (length less than $\eta$ times the median length, set to 0.6) in each bundle. We then reapplied hierarchical clustering.
